<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Cross-validation}
-->

# Cross-validation

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Comparing cross-validation methods

```{r}
library(data.table)
n.segments <- 10
seg.mean.vec <- 1:n.segments
data.per.segment <- 10
data.mean.vec <- rep(seg.mean.vec, each=data.per.segment)
n.data <- length(data.mean.vec)
n.validation.sets <- 100
n.folds.vec <- c(10, 2)
prop.valid.vec <- 1/n.folds.vec
sim.result.list <- list()
for(data.seed in 1:100){
  set.seed(data.seed)
  data.vec <- rnorm(n.data, data.mean.vec, 0.1)
  is.valid.vec.list <- list()
  for(n.folds in n.folds.vec){
    uniq.folds <- 1:n.folds
    n.seeds <- n.validation.sets/n.folds
    split.type <- sprintf("%d-fold %d times", n.folds, n.seeds)
    for(seed in 1:n.seeds){
      set.seed(seed)
      fold.vec <- sample(rep(uniq.folds, l=n.data))
      for(valid.fold in uniq.folds){
        is.valid.vec.list[[split.type]][[paste(seed, valid.fold)]] <-
          fold.vec==valid.fold
      }
    }
  }
  for(prop.valid in prop.valid.vec){
    split.type <- sprintf("%d%% %d times", 100*prop.valid, n.validation.sets)
    prop.vec <- c(subtrain=1-prop.valid, validation=prop.valid)
    for(split.i in 1:n.validation.sets){
      set.seed(split.i)
      is.valid.vec.list[[split.type]][[split.i]] <- binsegRcpp::random_set_vec(
        n.data, prop.vec) == "validation"
    }
  }
  loss.dt <- CJ(split.i=1:n.validation.sets, type=names(is.valid.vec.list))[, {
    is.valid <- is.valid.vec.list[[type]][[split.i]]
    bs.model <- binsegRcpp::binseg_normal(data.vec, is.validation.vec=is.valid)
    bs.model$splits[, data.table(
      segments,
      validation.loss)]
  }, by=.(split.i, type)]
  loss.stats <- loss.dt[, .(
    mean.valid.loss=mean(validation.loss)
  ), by=.(type, segments)]
  select.each.split <- loss.dt[
  , .SD[which.min(validation.loss)],
    by=.(type, split.i)]
  selected.times <- select.each.split[, .(
    times=.N
  ), by=.(type, segments)]
  selected.segments <- rbind(
    select.each.split[, .(
      selected=min(segments)
    ), by=.(method=paste(type, "min err, min segs"))],
    selected.times[, .(
      selected=segments[which.max(times)]
    ), by=.(method=paste(type, "min err, max times"))],
    loss.stats[, .(
      selected=segments[which.min(mean.valid.loss)]
    ), by=.(method=paste(type, "mean err, min err"))]
  )
  sim.result.list[[data.seed]] <- data.table(
    data.seed, selected.segments, n.segments)
}
sim.result <- do.call(rbind, sim.result.list)
(sim.err <- sim.result[, .(
  zero.one.loss=sum(selected != n.segments),
  L1.loss=sum(abs(selected-n.segments)),
  L2.loss=sum((selected-n.segments)^2)
), by=method][order(zero.one.loss)])
plot(data.vec)
```

The code above compares several types of cross-validation for
selecting the number of segments in simulated random normal data. The
table above shows various error rates which compare the selected
number of segments to the true number of segments in the
simulation. The best methods appears to be the ones which use min err,
max times.

## How many times is sufficient?

```{r}
n.segments <- 20
seg.mean.vec <- 1:n.segments
data.per.segment <- 5
data.mean.vec <- rep(seg.mean.vec, each=data.per.segment)
n.data <- length(data.mean.vec)
n.validation.sets <- 200
prop.valid <- c(0.01, 0.05, 0.1, 0.25, 0.5)
sim.result <- data.table(data.seed=1:100)[, {
  set.seed(data.seed)
  data.vec <- rnorm(n.data, data.mean.vec, 0.1)
  select.each.split <- CJ(split.i=1:n.validation.sets, prop.valid)[, {
    set.seed(split.i)
    prop.sets <- c(subtrain=1-prop.valid, validation=prop.valid)
    is.valid <- binsegRcpp::random_set_vec(
      n.data, prop.sets)=="validation"
    bs.model <- binsegRcpp::binseg_normal(
      data.vec, is.validation.vec=is.valid)
    bs.model$splits[, .(selected=segments[which.min(validation.loss)])]
  }, by=.(split.i, prop.valid)]
  data.table(n.splits=1:n.validation.sets)[, {
    select.each.split[split.i <= n.splits, .(
      times=.N
    ),
    by=.(prop.valid, selected)
    ][, .SD[which.max(times), .(selected)], by=prop.valid]
  }, by=n.splits]
}, by=data.seed]

if(require(ggplot2)){
  ggplot()+
    scale_color_gradient(low="red", high="black")+
    geom_line(aes(
      n.splits, selected,
      group=paste(data.seed, prop.valid),
      color=prop.valid),
      data=sim.result)+
    scale_y_continuous(breaks=seq(0, 100, by=10))
}

accuracy.dt <- sim.result[, .(
  correct=sum(selected==n.segments)
), by=.(prop.valid, n.splits)]
if(require(ggplot2)){
  gg <- ggplot()+
    geom_line(aes(
      n.splits, correct,
      group=prop.valid,
      color=prop.valid),
      size=2,
      data=accuracy.dt)+
    scale_y_continuous("number of correctly chosen data sets")+
    scale_color_gradient(low="red", high="black")
  if(require(directlabels)){
    direct.label(gg, "right.polygons")
  }else{
    gg
  }
}
```

The plot above suggests that 100 validation sets is sufficient.
